Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2016-10-10 forgerard.tex
responding to Greg's comments

\begin{abstract}
\color{red}
An empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and two independent color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence.
The two parameters exhibit the behavior expected from the two-parameter
\citet{1999PASP..111...63F} dust model with a per-supernova scatter in
total-to-selective extinction, an effective
%-----
$R^F_V=2.43^{+0.16}_{-0.17}$
%------
for the full sample, and a difference in the average $R^F_V$ between extremely blue and red supernovae .  The derived dust law is able to reproduce the optical colors of the highly reddened SN~2014J.
Extending the model to allow three color parameters gives a detection of a third parameter at  $>98$\% confidence.
\color{black}
\end{abstract}

@@

\section{Introduction}
Type~Ia supernovae (SNe~Ia) form a homogenous set of exploding stars and as such were early recognized and utilized as a powerful distance indicator 
and probe of cosmology \citep[e.g.][]{1992ARA&A..30..359B, 1993ApJ...415....1S}.  After further careful consideration of supernova data, it was recognized
that SN~Ia light-curve shapes \citep{1993ApJ...413L.105P} and colors \citep{1998A&A...331..815T} exhibit subtle signs of heterogeneity,
which being correlated with absolute magnitude
need to be considered when inferring distances.  Empirical models parameterizing SNe~Ia by their light-curve shape
and color were developed \citep{1996ApJ...473...88R, 1999ApJ...517..565P} that enabled absolute magnitude corrections
and accurate distance measurements of cosmological supernovae,
which 
were subsequently used in the discovery of the accelerating expansion of the Universe \citep{1998AJ....116.1009R,1999ApJ...517..565P}.

The two most commonly used supernova-cosmology light-curve fitters today are SALT2 \citep{2007A&A...466...11G} and MLCS2k2
\citep{2007ApJ...659..122J}.\footnote{Light-curve fitters with more flexible degrees of freedom
\citep[e.g.][]{2008ApJ...681..482C, 2011AJ....141...19B, 2011ApJ...731..120M} are available and have for
the most part been used to study SN~Ia heterogeneity.}
They remain two-parameter models, with one parameter characterizing light-curve shape and the other
 color.  The physical cause of the color diversity is interpreted differently by the two sets of authors: 
\citet{2007A&A...466...11G} pragmatically extract color variation empirically from SNe that span a wide range of colors, with no attribution
to either intrinsic or extrinsic origins;
\citet{2007ApJ...659..122J}
attribute changes in color to the attenuation of light from host-galaxy dust. Neither model accommodates the naive expectation of both independent
intrinsic and extrinsic color variability.

There is evidence that supports the expectation that a single parameter cannot describe the full range
of colors seen in the SN~Ia population.  One approach to look for color diversity is to find correlations between color and spectral features.
\citet{2009ApJ...699L.139W, 2011ApJ...729...55F} find two subpopulations distinguished
by Si velocity with differing $B_{max}-V_{max}$.
\citet{2015MNRAS.451.1973S}
find that high-velocity SiII$\lambda$6355 is found in objects that have red ultraviolet/optical colors near maximum brightness.

\textcolor{red}{
Another approach to probe color diversity is through multiple colors (at least 3 bands)
of a single supernova.  Colors ratios are sensitive to processes of the responsible physics.   For example,
relative dust absorption varies as a function of wavelength depending on grain size and shape,
independent (to first order) of the amount of dust along the line of sight.
}
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} find wide
ranges of total-to-selective extinction with average values significantly lower than $R_V = 3.1$,
the canonical value for diffuse Milky Way dust.
They also confirm the \citet{2011ApJ...729...55F} finding that low $R_V$ is associated with high-extinction supernovae.
In contrast, \citet{2011A&A...529L...4C} argue that after accounting for the diversity of spectral features,
the standard $R_V=3.1$ is recovered on average
\textcolor{red}{though only after excluding their $>3 \sigma$ reddest supernovae.}

\textcolor{red}
{
Hierarchical modeling has recently opened
the study of intrinsic supernova color based on SN~Ia Hubble diagrams.
\citet{2016arXiv160904470M} find that the distribution of
observed supernova magnitudes is better explained by a combination of intrinsic color dispersion and
$R_V=3.7$ dust, over dust with no color dispersion.
They draw these conclusions by using only one
derived ``color'', the SALT2 fit parameter.
}

The Nearby Supernova Factory \citep[SNfactory;][]{2004SPIE.5249..146L} has collected  
spectrophotometric time series of hundreds of Hubble-flow $0.03<z<0.08$ SNe~Ia.   The $3200$--$10000$~\AA\ spectral coverage
provides measurements of an array of supernova spectral features while also providing synthetic broadband photometry
spanning near-UV to near-IR SN-frame wavelengths.  Well over a hundred of these supernovae have temporal coverage over
peak brightness.  This dataset provides a homogenous sample with which to study SN~Ia colors and spectral features together.

In this article we use the idea that spectral indicators carry information on intrinsic supernova colors at peak magnitude.
We allow for two independent color parameters, one due to intrinsic supernova
properties and the other due to
extrinsic processes, which we attribute to dust.  The data used in this analysis are described in \S\ref{data:sec}.  Our supernova model is presented in
\S\ref{model:sec} and results are given in \S\ref{results:sec}.  In \S\ref{sn2014j:sec} our results are applied to the out-of-sample
supernova SN2014J that has been previously identified as having extreme total-to-selective extinction.
We conclude with a discussion of our findings in \S\ref{discussion:sec}.

@@

Our analysis uses the spectrophotometric data set obtained by
the SNfactory with the SuperNova Integral Field
Spectrograph \citep[SNIFS,][]{2004SPIE.5249..146L}.  SNIFS is a fully integrated
instrument optimized for automated observation of point sources on a
structured background over the full ground-based optical window at
moderate spectral resolution ($R \sim 500$).  It consists of a
high-throughput wide-band pure-lenslet integral field spectrograph
\citep[IFS, ``\`a la TIGER;''][]{1995A&AS..113..347B,2000ASPC..195..173B,2001MNRAS.326...23B}, a
multi-filter photometric channel to image the field in the vicinity of
the IFS for atmospheric transmission monitoring simultaneous with
spectroscopy, and an acquisition/guiding channel.  The IFS possesses a
fully-filled $6\farcs 4 \times 6\farcs 4$ spectroscopic field of view
subdivided into a grid of $15 \times 15$ spatial elements, a
dual-channel spectrograph covering 3200--5200~\AA\ and 5100--10000~\AA\
simultaneously, and an internal calibration unit (continuum and arc
lamps).  SNIFS is mounted on the south bent Cassegrain port of the
University of Hawaii 2.2~m telescope on Mauna Kea, and is operated
remotely.  Observations are reduced using the SNfactory's dedicated data
reduction pipeline, similar to that presented in \S4 of \citet{2001MNRAS.326...23B}.
A discussion of the software pipeline is presented in
\citet{2006ApJ...650..510A} and is updated in \citet{2010ApJ...713.1073S}.  A detailed
description of host-galaxy subtraction is given in \citet{2011MNRAS.418..258B}.

The current status of the data and its treatment is presented in \citet{2015ApJ...815...58F}.
We provide a brief summary of the points important for this analysis.
The spectral time-series  are corrected for Milky Way dust
extinction \citep{1989ApJ...345..245C,1998ApJ...500..525S} but not for the
effects of circumstellar or host-galaxy dust extinction.  
Each spectral time series is
blue-shifted to rest-frame, and the fluxes are converted to luminosity assuming
distances expected for the supernova redshifts given a flat
$\Lambda$CDM cosmology with $\Omega_M = 0.28$ (with an arbitrarily selected
$H_0$ since the current analysis does not depend on the absolute magnitude scale).

Supernova-frame synthetic photometry is generated for a top-hat filter system
comprised of five 
\color{red}
bands with the following wavelength ranges: $U$ $[3300.00 - 3978.02]$\AA;
$B$ $[3978.02-4795.35]$\AA;
$V$ $[4795.35-5780.60]$\AA;
$R$ $[5780.60-6968.29]$\AA;
$I$ $[6968.29-8400.00]$\AA.
\color{black}
For each supernova, the magnitudes at peak brightness are determined using single-band SALT2 fits.
The equivalent widths of SiII$\lambda 4141$ and the CaII H\&K features are computed as
in \citet{2008A&A...477..717B} and the 
\textcolor{red}{wavelength of the SiII$\lambda 6355$ feature}
as in \citet{chotard:thesis}.
Equivalent widths and the
\textcolor{red}{SiII$\lambda 6355$ wavelength are taken from spectra  within $\pm 2.5$ days from maximum, in the case of multiple spectra the one
closest to maximum is used \citep{chotard:thesis}.
The weighted mean is used for the  SiII$\lambda 6355$ wavelength.}

Our analysis sample is comprised by the
166
supernovae that have the data coverage to 
give photometric and spectroscopic statistics described above.
The data are given in Table~\ref{data:tab}.
The peak magnitude uncertainties do have covariance, which are accounted
for in the analysis; only the standard deviation is included in the table.

@@

\tablecaption{Supernova Spectral-Feature and Peak-Magnitude Data\label{data:tab}}

@@

Absolute photometric calibration is of extreme importance for the SNfactory.  This analysis, however, is confined to the relative colors for a low-redshift set of supernovae
and is thus insensitive to uncertainties in these zeropoints.

\section{Supernova Model: Two Extrinsic Parameters}
\label{model:sec}

We hypothesize that at peak brightness
SN~Ia broadband magnitudes and colors are correlated with
spectral features: equivalent widths and line positions are considered as statistics localized in wavelength that are insensitive to variations in
broadband colors.
%Spectra may not efficiently capture all variations in broadband magnitude, so an additional source
%of intrinsic variation is accommodated.
The intrinsic spectral  parameters may not deterministically predict magnitudes, but rather do so with some intrinsic dispersion.
\textcolor{red}{The intrinsic magnitudes are then
modified by an extrinsic physical process (i.e.\ dust) to produce apparent magnitudes.}
\subsection{Model}
We assume 
that  peak intrinsic $UBVRI$ magnitudes are linearly dependent
on the
 equivalent widths of the CaII H\&K and SiII$\lambda$4141 spectral features
$EW_{Ca}$, $EW_{Si}$,
\textcolor{red}{ and $\lambda_{Si}$ the wavelength of the minimum of 
the SiII$\lambda6355$ feature}:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F}.
\textcolor{red}{
The explicit omission of light-curve shape in our model is compensated by its proxy
$EW_{Si}$ at peak brightness}
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
% An intrinsic-color parameter $D$ describes  intrinsic magnitude/color variation unaccounted for by the spectral features.
Residual dispersion is described by a Normal distribution with a parameterized covariance matrix
$C_c$.  A grey magnitude offset $\Delta$ is included for each supernova
to capture intrinsic dispersion and peculiar-velocity errors introduced when converting fluxes to luminosity.
The observed magnitudes are linearly dependent on the
extrinsic-color parameters $k_0$ and $k_1$.  The observables
$U_o, B_o, V_o, R_o, I_o$, $EW_{Ca,o}$, $EW_{Si,o}$, $\lambda_{Si,o}$
shown in Table~\ref{data:tab} have Gaussian measurement uncertainty with covariance $C$.
The model is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_0+\alpha_0 EW_{Ca} + \beta_0 EW_{Si} + \eta_0 \lambda_{Si} \\
c_1+\alpha_1 EW_{Ca} + \beta_1 EW_{Si} + \eta_1 \lambda_{Si}  \\
c_2+\alpha_2 EW_{Ca} + \beta_2 EW_{Si} + \eta_2 \lambda_{Si} \\
c_3+\alpha_3 EW_{Ca} + \beta_3 EW_{Si} + \eta_3 \lambda_{Si} \\
c_4+\alpha_4 EW_{Ca} + \beta_4 EW_{Si}+ \eta_4 \lambda_{Si}
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\B +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\
V+\gamma^0_{2} k_0+\gamma^1_{2} k_1\\R+\gamma^0_{3} k_0 + \gamma^1_{3} k_1\\I+\gamma^0_{4} k_0+\gamma^1_{4} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
The global parameters $c$, $\alpha$, $\beta$,  and $\eta$,  are the intercepts and slopes of the linear relationships that
relate intrinsic magnitudes with the spectral-feature parameters.
(In this article we freely interchange the parameter indices with  $01234$ or $UBVRI$.)
The global parameters $\gamma^0$, $\gamma^1$  are the slopes that connect the extrinsic-color
parameters to observed magnitudes.

To constrain the degrees of freedom and degeneracies inherent in the model we impose that
\color{red}
each of the vectors $k_0+1/N$ and $k_1+1/N$ are simplexes, where $N$ is the number of objects in our sample: these
conditions break the 
degeneracies  $\gamma ^0\rightarrow a\gamma^0$, $k_0 \rightarrow a^{-1}k_0$ and $\gamma^1 \rightarrow a\gamma^1$, $k_1 \rightarrow a^{-1}k_1$
up to the $\pm$ sign.
Furthermore we impose
\color{black} 
\begin{equation}
\langle \Delta \rangle=0, \langle k_0 \rangle=0, \langle k_1 \rangle=0, \gamma^0_0 > 0, \gamma^1_0 < 0.
\end{equation}
The first two specify the definition of zero color relative to which the color excess is measured.    The latter two exclude degenerate posterior space
associated
with  simultaneous sign flips of
$\gamma^0$--$k_0$ and $\gamma^1$--$k_1$
\color{red}
(when running our fits, $k_0$ and $k_1$ are set to
identical initial conditions, so the difference in the signs of $\gamma^0_0$ and $\gamma^1_0$ distinguishes the two.)
As  $\gamma^0$--$k_0$ and $\gamma^1$--$k_1$ are degenerate with
each other,
the different initial and boundary conditions for $\gamma^0_0$ and $\gamma^1_0$ break the degeneracy of our mixture model;
otherwise identical initial conditions for $\gamma^0$, $\gamma^1$ produce indistinguishable posteriors
with relatively broad credible intervals and lower maximum likelihood.

@@

Having the intrinsic dispersion $C_c$ as fit parameters seemingly introduces degeneracy in the model, as magnitude and color variation
ascribed to $\Delta$, $\gamma^0 k_0$, and $\gamma^1 k_1$ could also be attributed to intrinsic dispersion.  There are several features of the model
that drives the assignation of variations away from $C_c$:  Maximizing the posterior disfavors the increase of $\det{(C_c)}$;
The distributions of $\Delta$, $k_0$, and $k_1$ turn out to
be non-Gaussian, and so are not well described by a Normal covariance; Grey offsets would appear as a constant
in all elements of the covariance matrix, which is disfavored for the Bayesian prior selected for $C_c$.

In a Bayesian analysis the priors must be described.  A flat prior is used for all parameters except
for the covariance matrix $C_c$, which is constructed from a correlation matrix with  $\nu=4$  LKJ prior\footnote{
Visualization of the LKJ correlation distribution can be found in \url{http://www.psychstatistics.com/2014/12/27/d-lkj-priors/}.}
\citep{Lewandowski20091989} and standard
deviations $\sigma_i = \sqrt{C_{c,ii}}$ with a  Cauchy distribution prior with location
 $0.1$ and scale $0.1$ mag restricted to positive values.  (We find that imposing a stricter assumption of a
 diagonal covariance matrix, with no reference to a correlation matrix with LKJ prior, yields the same parameter distributions within
 uncertainty).

>>>>

\begin{abstract}
\color{red}
An empirical model for SN~Ia peak magnitudes with two color parameters and dependence on the equivalent widths of CaII, SiII, and SiII velocity
is applied to the supernova sample of the Nearby Supernova Factory.  The peak magnitudes in synthetic
broadband photometry and their colors are found to be 
dependent on the spectral equivalent widths and two independent color parameters, to better than 
%-----
$(1-5\times10^{-5})$
%------
confidence.
The two parameters exhibit the behavior expected from the two-parameter
\citet{1999PASP..111...63F} dust model with a per-supernova scatter in
total-to-selective extinction, an effective
%-----
$R^F_V=2.43^{+0.16}_{-0.17}$
%------
for the full sample, and a difference in the average $R^F_V$ between extremely blue and red supernovae .  The derived dust law is able to reproduce the optical colors of the highly reddened SN~2014J better than the Fitzpatrick dust model.
Extending the model to allow three color parameters gives almost identical results for
the two parameters, plus a detection of a third parameter at  $>98$\% confidence.
\color{black}
\end{abstract}

@@

\section{Introduction}
Type~Ia supernovae (SNe~Ia) form a homogenous set of exploding stars and as such were early recognized and utilized as a powerful distance indicator 
and probe of cosmology \citep[e.g.][]{1992ARA&A..30..359B, 1993ApJ...415....1S}.  After further careful consideration of supernova data, it was recognized
that SN~Ia light-curve shapes \citep{1993ApJ...413L.105P} and colors \citep{1996ApJ...473...88R, 1998A&A...331..815T} exhibit subtle signs of heterogeneity,
which being correlated with absolute magnitude
need to be considered when inferring distances.  Empirical models parameterizing SNe~Ia by their light-curve shape \citep{1996ApJ...473...88R,
1997ApJ...483..565P,
1999ApJ...517..565P}
and color  \citep{1996ApJ...473...88R}  were developed that enabled absolute magnitude corrections
and accurate distance measurements of cosmological supernovae,
which 
were subsequently used in the discovery of the accelerating expansion of the Universe \citep{1998AJ....116.1009R,1999ApJ...517..565P}.

The two most commonly used supernova-cosmology light-curve fitters today are SALT2 \citep{2007A&A...466...11G} and MLCS2k2
\citep{2007ApJ...659..122J}.\footnote{Light-curve fitters with more flexible degrees of freedom
\citep[e.g.][]{2008ApJ...681..482C, 2011AJ....141...19B, 2011ApJ...731..120M} are available and have for
the most part been used to study SN~Ia heterogeneity.}
They remain two-parameter models, with one parameter characterizing light-curve shape and the other
 color.
 \textcolor{red}{In SALT2 the light curve shapes are described by a multiplicative scaling (``stretch'')  of the time evolution,
 whereas MLCS2k2 varies shape with additive corrections to magnitudes.}
 The physical cause of the color diversity is interpreted differently by the two sets of authors: 
\citet{2007A&A...466...11G} pragmatically extract color variation empirically from SNe that span a wide range of colors, with no attribution
to either intrinsic or extrinsic origins;
\citet{2007ApJ...659..122J}
attribute changes in color
\textcolor{red}{partially to intrinsic variations connected to light-curve shape, and partially}
to the attenuation of light from host-galaxy dust.

There is evidence that supports the expectation that a single parameter beyond light-curve shape  cannot describe the full range
of colors seen in the SN~Ia population.  One approach to look for color diversity is to find correlations between color and spectral features.
\citet{2009ApJ...699L.139W, 2011ApJ...729...55F} find two subpopulations distinguished
by Si velocity with differing $B_{max}-V_{max}$; this color correlation, \textcolor{red}{ in addition to one with $B-R$, is confirmed by
\citet{2014ApJ...797...75M}.
}
\citet{2015MNRAS.451.1973S}
find that high-velocity SiII$\lambda$6355 is found in objects that have red ultraviolet/optical colors near maximum brightness.
\textcolor{red}{
\citet{2011MNRAS.413.3075M} argue that viewing angle is an important determinant in controlling supernova color, which is
traced by wavelength shifts in nebular emission lines; such correlations are also seen by \citet{2011A&A...534L..15C}.
}

\textcolor{red}{
Another approach to probe color diversity is through multiple colors (at least 3 bands)
of a single supernova.  Colors ratios are sensitive to processes of the responsible physics.   For example,
relative dust absorption varies as a function of wavelength depending on grain size and shape,
independent (to first order) of the amount of dust along the line of sight.
\citet{2013ApJ...779...23M} find diversity in SNe~Ia UV emission, which they use to distinguish subclasses based on UV-optical colors.}
\textcolor{red}{
Such studies measuring $R_V$ are being advanced with the development of flexible empirical light curve models that accommodate flexibility in multi-band colors
\citep[e.g.][]{2011ApJ...731..120M}.}
\citet{2014ApJ...789...32B, 2015MNRAS.453.3300A} find wide
ranges of total-to-selective extinction with average values significantly lower than $R_V = 3.1$,
the canonical value for diffuse Milky Way dust.
They also confirm the \citet{2011ApJ...731..120M, 2011ApJ...729...55F} finding that low $R_V$ is associated with high-extinction supernovae.
In contrast, \citet{2011A&A...529L...4C} argue that after accounting for the diversity of spectral features,
the standard $R_V=3.1$ is recovered on average.

\textcolor{red}
{
Hierarchical modeling has recently opened
the study of intrinsic supernova color based on SN~Ia Hubble diagrams; latent parameters that are not directly tied to observables,
such as multiple parameters that simultaneously influence color, can be determined.
\citet{2016arXiv160904470M} take the approach of modeling the distribution of their parameters, to find that the distribution of
observed supernova magnitudes is better explained by a combination of intrinsic color dispersion and
$R_B=3.7$ dust, over dust with no color dispersion.
They draw these conclusions by using only the SALT2 fit $X_1$ parameter as the summary statistic that describes color.
}

The Nearby Supernova Factory \citep[SNfactory;][]{2002SPIE.4836...61A} has systematically observed the
spectrophotometric time series of hundreds of Hubble-flow $0.03<z<0.08$ SNe~Ia.   The $3200$--$10000$~\AA\ spectral coverage
provides measurements of an array of supernova spectral features while also providing synthetic broadband photometry
spanning near-UV to near-IR SN-frame wavelengths.  \textcolor{red}{SNfactory specifically targeted objects
early in their temporal evolution, so} that well over a hundred of these supernovae have  coverage over
peak brightness.  This dataset provides a homogenous sample with which to study SN~Ia colors and spectral features together.

In this article we use the idea that spectral indicators carry information on intrinsic supernova colors at peak magnitude.
We allow for \textcolor{red}{three independent color parameters, one attributed to intrinsic supernova
properties and two  attributed} to
extrinsic processes, which we attribute to dust.  The data used in this analysis are described in \S\ref{data:sec}.  \textcolor{red}{A
first analysis omitting the intrinsic parameter is presented in
\S\ref{model:sec}, where we show that the properties of the two extrinsic parameters
are consistent with the expectations of Milky Way-parameterized dust, and do an excellent job of fitting the data of the out-of-sample
supernova SN2014J that has been previously identified as having extreme total-to-selective extinction.
In \S\ref{model2:sec} we add to analysis the third intrinsic parameter, and find strong evidence for its influence on observed magnitudes.
We summarize our findings, and relate our results with light-curve shape, SiII velocity, and the ``mass step'' in SN~Ia Hubble
residuals in \S\ref{discussion:sec}.}

@@

Our analysis uses the spectrophotometric data set obtained by
the SNfactory with the SuperNova Integral Field
Spectrograph \citep[SNIFS,][]{2004SPIE.5249..146L}.  SNIFS is a fully integrated
instrument optimized for automated observation of point sources on a
structured background over the full ground-based optical window at
moderate spectral resolution ($R \sim 500$).  It consists of a
high-throughput wide-band lenslet integral field spectrograph, a
multi-filter photometric channel to image the field in the vicinity of
the IFS for atmospheric transmission monitoring simultaneous with
spectroscopy, and an acquisition/guiding channel.  The IFS possesses a
fully-filled $6\farcs 4 \times 6\farcs 4$ spectroscopic field of view
subdivided into a grid of $15 \times 15$ spatial elements, a
dual-channel spectrograph covering 3200--5200~\AA\ and 5100--10000~\AA\
simultaneously, and an internal calibration unit (continuum and arc
lamps).  SNIFS is mounted on the south bent Cassegrain port of the
University of Hawaii 2.2~m telescope on Mauna Kea, and is operated
remotely.  Observations are reduced using the SNfactory's dedicated data
reduction pipeline, similar to that presented in \S4 of \citet{2001MNRAS.326...23B}.
A discussion of the software pipeline is presented in
\citet{2006ApJ...650..510A} and is updated in \citet{2010ApJ...713.1073S}. 
The flux calibration is presented in \citet{2013A&A...549A...8B}.  A detailed
description of host-galaxy subtraction is given in \citet{2011MNRAS.418..258B}.

A recent description of the data is presented in \citet{2015ApJ...815...58F}.
We provide a brief summary of the points important for this analysis.
The spectral time-series  are corrected for Milky Way dust
extinction \citep{1989ApJ...345..245C,1998ApJ...500..525S}.  
Each spectral time series is
blue-shifted to rest-frame
based on the systemic redshift of the host \citep[c.f.][]{2013ApJ...770..107C}, and the fluxes are converted to luminosity assuming
distances expected for the supernova redshifts given a flat
$\Lambda$CDM cosmology with $\Omega_M = 0.28$ (with an arbitrarily selected
$H_0$ since the current analysis does not depend on the absolute magnitude scale).

Supernova-frame synthetic supernova-frame photometry is generated for a top-hat filter system
comprised of five 
\color{red}
bands with the following wavelength ranges: $U$ $[3300.00 - 3978.02]$\AA;
$B$ $[3978.02-4795.35]$\AA;
$V$ $[4795.35-5780.60]$\AA;
$R$ $[5780.60-6968.29]$\AA;
$I$ $[6968.29-8400.00]$\AA.
For each supernova, the magnitudes at peak brightness are regressed from data within 5-days of peak brightness
using single-band SALT2 fits.
\color{black}
The equivalent widths of SiII$\lambda 4141$ and the CaII H\&K features are computed as
in \citet{2008A&A...477..717B} and the 
\textcolor{red}{wavelength of the SiII$\lambda 6355$ feature}
as in \citet{chotard:thesis}.
Equivalent widths and the
\textcolor{red}{SiII$\lambda 6355$ wavelength are taken from spectra  within $\pm 2.5$ days from maximum.}

Our analysis sample is comprised by the
166
supernovae that have the data coverage to 
give photometric and spectroscopic statistics described above.
The data are given in Table~\ref{data:tab}.
The peak magnitude uncertainties do have covariance, which is accounted
for in the analysis; only the standard deviation is included in the table.

@@

\tablecaption{Supernova Spectral-Feature and Peak-Magnitude Data \textcolor{red}{TABLE MAY GO IF THERE IS A NICO PAPER}
\label{data:tab}}

@@

%Absolute photometric calibration is of extreme importance for the supernova cosmology.
%This analysis, however, is confined to the relative colors for a low-redshift set of supernovae
%and is thus insensitive to uncertainties in these zeropoints.

\section{Supernova Model I: Two Extrinsic Parameters}
\label{model:sec}

We hypothesize that at peak brightness
SN~Ia broadband magnitudes and colors are correlated with
spectral features: equivalent widths and line positions are considered as statistics localized in wavelength that are insensitive to variations in
broadband colors.
%Spectra may not efficiently capture all variations in broadband magnitude, so an additional source
%of intrinsic variation is accommodated.
These intrinsic spectral  parameters may not deterministically predict magnitudes, but rather do so with some intrinsic dispersion.
\textcolor{red}{The intrinsic magnitudes are then
modified by an extrinsic physical process (i.e.\ dust) to produce apparent magnitudes.}
\subsection{Model}
We assume 
that  peak intrinsic $UBVRI$ magnitudes are linearly dependent
on the
 equivalent widths of the CaII H\&K and SiII$\lambda$4141 spectral features
$EW_{Ca}$, $EW_{Si}$,
\textcolor{red}{ and $\lambda_{Si}$ the wavelength of the minimum of 
the SiII$\lambda6355$ feature}:
these spectral features are associated with SN~Ia  spectroscopic diversity  
\citep{2006PASP..118..560B, 2008A&A...492..535A, 2009A&A...500L..17B, 2009PASP..121..238B, 2009ApJ...699L.139W, 2011ApJ...729...55F}.
\textcolor{red}{
The explicit omission of light-curve shape in our model is compensated by its proxy
$EW_{Si}$ at peak brightness}
\citep{2008A&A...492..535A, 2011A&A...529L...4C}. 
% An intrinsic-color parameter $D$ describes  intrinsic magnitude/color variation unaccounted for by the spectral features.
Residual dispersion is described by a Normal distribution with a parameterized covariance matrix
$C_c$.  A grey magnitude offset, $\Delta$, is included for each supernova
to capture intrinsic dispersion and peculiar-velocity errors introduced when converting fluxes to luminosity.
The observed magnitudes are linearly dependent on the
extrinsic-color parameters $k_0$ and $k_1$.  The observables
$U_o, B_o, V_o, R_o, I_o$, $EW_{Ca,o}$, $EW_{Si,o}$, $\lambda_{Si,o}$
shown in Table~\ref{data:tab} have Gaussian measurement uncertainty with covariance $C$.
The model is written as
\begin{equation}
\begin{pmatrix}
U\\B\\V\\R\\I
\end{pmatrix}
\sim \mathcal{N}
\left(
\Delta +
\begin{pmatrix}
c_0+\alpha_0 EW_{Ca} + \beta_0 EW_{Si} + \eta_0 \lambda_{Si} \\
c_1+\alpha_1 EW_{Ca} + \beta_1 EW_{Si} + \eta_1 \lambda_{Si}  \\
c_2+\alpha_2 EW_{Ca} + \beta_2 EW_{Si} + \eta_2 \lambda_{Si} \\
c_3+\alpha_3 EW_{Ca} + \beta_3 EW_{Si} + \eta_3 \lambda_{Si} \\
c_4+\alpha_4 EW_{Ca} + \beta_4 EW_{Si}+ \eta_4 \lambda_{Si}
\end{pmatrix}
,C_{c}
\right)
\label{ewsiv:eqn}
\end{equation}
\begin{equation}
\begin{pmatrix}
U_o\\B_o\\ V_o\\R_o\\I_o\\EW_{Si, o}\\ EW_{Ca, o} \\ \lambda_{Si, o}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
U +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\B +\gamma^0_{1} k_0 +\gamma^1_{1} k_1 \\
V+\gamma^0_{2} k_0+\gamma^1_{2} k_1\\R+\gamma^0_{3} k_0 + \gamma^1_{3} k_1\\I+\gamma^0_{4} k_0+\gamma^1_{4} k_1\\
EW_{Si}\\ EW_{Ca} \\ \lambda_{Si}
\end{pmatrix}
,C
\right).
\label{dust:eqn}
\end{equation}
The global parameters $c$, $\alpha$, $\beta$,  and $\eta$,  are the intercepts and slopes of the linear relationships that
relate intrinsic magnitudes with the spectral-feature parameters.
(In this article we freely interchange the parameter indices with  $01234$ or $UBVRI$.)
The global parameters $\gamma^0$, $\gamma^1$  are the slopes that connect the extrinsic-color
parameters to observed magnitudes.

To constrain the degrees of freedom and degeneracies inherent in the model we impose that
\color{red}
each of the vectors $k_0+1/N$ and $k_1+1/N$ are simplexes, where $N$ is the number of objects in our sample: these
conditions break the 
degeneracies  $\gamma ^0\rightarrow a\gamma^0$, $k_0 \rightarrow a^{-1}k_0$ and $\gamma^1 \rightarrow a\gamma^1$, $k_1 \rightarrow a^{-1}k_1$
up to the $\pm$ sign.
Furthermore we impose
\color{black} 
\begin{equation}
\langle \Delta \rangle=0, \langle k_0 \rangle=0, \langle k_1 \rangle=0, \gamma^0_0 > 0, \gamma^1_0 < 0.
\end{equation}
The first two conditions specify the definition of zero color relative to which the color excess is measured:
in a linear model this value is arbitrary.    The latter two exclude degenerate posterior space
associated
with  simultaneous sign flips of
$\gamma^0$--$k_0$ and $\gamma^1$--$k_1$
\color{red}
(when running our fits, $k_0$ and $k_1$ are set to
identical initial conditions, so the difference in the signs of $\gamma^0_0$ and $\gamma^1_0$ distinguishes the two.)
As  $\gamma^0$--$k_0$ and $\gamma^1$--$k_1$ are degenerate with
each other,
the different initial and boundary conditions for $\gamma^0_0$ and $\gamma^1_0$ break the degeneracy of our mixture model;
otherwise identical initial conditions for $\gamma^0$, $\gamma^1$ produce indistinguishable posteriors
with relatively broad credible intervals and lower maximum likelihood.

@@

Having the intrinsic dispersion, $C_c$, as fit parameters seemingly introduces degeneracy in the model, as magnitude and color variation
ascribed to $\Delta$, $\gamma^0 k_0$, and $\gamma^1 k_1$ could also be attributed to intrinsic dispersion.  There are several features of the model
that drive the assignation of variations away from $C_c$:  Maximizing the posterior disfavors the increase of $\det{(C_c)}$;
The distributions of $\Delta$, $k_0$, and $k_1$ turn out to
be non-Gaussian, and so are not well described by the Normal covariance we impose in $C_c$; the grey magnitude offsets, $\Delta$, would appear as a constant
in all elements of the covariance matrix, which is disfavored for the Bayesian prior selected for $C_c$.

In a Bayesian analysis such as this the priors must be described.  A flat prior is used for all parameters except
for the covariance matrix $C_c$, which is constructed from a correlation matrix with  $\nu=4$  LKJ prior\footnote{
Visualization of the LKJ correlation distribution can be found in \url{http://www.psychstatistics.com/2014/12/27/d-lkj-priors/}.}
\citep{Lewandowski20091989} and standard
deviations $\sigma_i = \sqrt{C_{c,ii}}$ with a  Cauchy distribution prior with location
 $0.1$ and scale $0.1$ mag restricted to positive values.
 \textcolor{red}{This prior for a covariance matrix is recommended by STAN.}
 (We find that imposing a stricter assumption of a
 diagonal covariance matrix, with no reference to a correlation matrix with LKJ prior, yields the same parameter distributions within
 uncertainty).
